---
title: "Sama Issa Basic inferential data analysis"
output:
  pdf_document: default
  html_document: default
---

## Overview

Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

1. Load the ToothGrowth data and perform some basic exploratory data analyses
2. Provide a basic summary of the data.
3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 
4. State your conclusions and the assumptions needed for your conclusions.

## inferential data analysis

### Load the ToothGrowth data and perform some basic exploratory data analyses                     
First we load the ToothGrowth dataset
```{r}
data(ToothGrowth)
```

### Provide a basic summary of the data                                
basic summary of data 
```{r}
str(ToothGrowth)
summary(ToothGrowth)

library(ggplot2)
g <- ggplot(ToothGrowth, aes(x = factor(dose), y = len))
g <- g + facet_grid(.~supp)
g <- g + geom_boxplot(aes(fill = supp))
g <- g + labs(title = "Tooth Length by Dosage for Every Supplement")
g <- g + labs(x = "Dose (mg/day)", y = "Tooth Length")
print(g)
```

### Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

```{r}
# Comparison by delivery method for the same dosage
t05 <- subset(ToothGrowth, dose == "0.5")
t1 <- subset(ToothGrowth, dose == "1")
t2 <- subset(ToothGrowth, dose == "2")

## Testing Dose wise
t05 <- t.test(len ~ supp, data = t05)
t1 <- t.test(len ~ supp, data = t1)
t2 <- t.test(len ~ supp, data = t2)

summaryBYsupp <- data.frame(
      "p-value" = c(t05$p.value, t1$p.value, t2$p.value),
      "Conf.Low" = c(t05$conf.int[1],t1$conf.int[1], t2$conf.int[1]),
      "Conf.High" = c(t05$conf.int[2],t1$conf.int[2], t2$conf.int[2]),
      row.names = c("Dosage .05","Dosage 1","Dosage 2"))

# Show the data table 
print("Dosage .05 :")
print(paste("    p-value :", t05$p.value))
print(paste("    Conf.Low :", t05$conf.int[1]))
print(paste("    Conf.High :", t05$conf.int[2]))

print("Dosage 1 :")
print(paste("    p-value :", t1$p.value))
print(paste("    Conf.Low :", t1$conf.int[1]))
print(paste("    Conf.High :", t1$conf.int[2]))

print("Dosage 2 :")
print(paste("    p-value :", t2$p.value))
print(paste("    Conf.Low :", t2$conf.int[1]))
print(paste("    Conf.High :", t2$conf.int[2]))

```

### State your conclusions and the assumptions needed for your conclusions                      


OJ delivers more tooth growth than VC for dosages 0.5 & 
1.0. OJ and VC deliver the same amount of tooth growth for 
dose amount 2.0 mg/day.  For the entire data set we cannot conclude OJ is more effective that VC.
